Machine Learning, Architectural Styles and Property Values

### 2020-04-13 text/manuscript_assa.Rmd
Tables and Figures on separate pages
Are we back in the game?

---
title: "Machine Learning, Architectural Styles and Property Values"
author: "Thies Lindenthal* (University of Cambridge) & Erik B. Johnson (University of Alabama)"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: Externalities of architecture, machine learning
fig_caption: yes
header-includes:
   - \usepackage{soul}
   - \usepackage{color}
   - \sethlcolor{yellow}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{graphicx}
   - \usepackage{grffile}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{multirow}
   - \doublespacing
   - \usepackage[backend=bibtex, style=authoryear]{biblatex}
fig_width: 6
fig_height: 4
output: 
  pdf_document:
    keep_tex: true
    number_sections: true
bibliography: references.bib
abstract: "This paper first introduces an algorithm that collects pictures of individual buildings from Google Street View. Second, it trains a deep convolutional neural network (CNN) to classify residential buildings into architectural styles, taking into account spatial dependencies of these styles. Third, it investigates whether architectural styles influence house prices. For re-sales, the architectural style is a significant determinant of transaction prices while no such effect is found for new buildings. Additionally, we are able to provide guidance on how to detect and overcome some of the limitations of machine learning methods through a large-scale comparison of predictions and expert classifications." 
---

# Introduction
\let\svthefootnote\thefootnote
\let\thefootnote\relax\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Peter Brummund, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}
\addtocounter{footnote}{-1}\let\thefootnote\svthefootnote
If new developments were only pleasing to the eye, so Britain's *Building Better, Building Beautiful Commission* advising Government, nimbyism would cease and housing supply could finally reach the levels demanded by a growing and more affluent population [@economistlineofbeauty2018]. The inaugural chair of this new commission boldly suggested that if we built *``as our Georgian and Victorian forebears built* [...]. *All objections to new building would slip away in the sheer relief of the public''* [@scruton2018]. Even Prince Charles, in similar spirit, put forward ten principles for urban growth and architecture that emphasize tradition and aesthetics [@princecharles2014]. When governments and princes occupy themselves with beauty, economists surely may take a closer look as well. In this paper, we aim to identify the architectural styles of residential buildings using computer vision techniques and to empirically search for any transaction price premia associated with these styles. If aesthetics were as strong a force in the built environment as claimed, we should indeed find a significant effect on property prices.

@@

Specifically, we first look up the nearest Google Street View panorama point (green dot in Figure \ref{fig:fan}) based on the centroid (red dot) coordinates of a given building obtained from Ordnance Survey maps. We then perform a viewshed analysis and identify which exterior walls are visible from the panorama point, ignoring any wall segments where the direct line of sight from the panorama point is obstructed by other buildings. We can then estimate the camera bearing (green line) and zoom factor, based on the fan of the lines of sight (in blue).\footnote{Code necessary to replicate this analysis is available from the author's GitHub.com code repositories: \href{https://github.com/thies}{https://github.com/thies}.}

@@

@@

\begin{table}[!htbp] \centering 
  \caption{Summary statistics residential property transactions} 
  \label{tab:sumstats} 
\begin{tabular}{@{\extracolsep{5pt}}lccccccc} 
\toprule
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Pctl(25)} & \multicolumn{1}{c}{Pctl(75)} & \multicolumn{1}{c}{Max} \\ 
\midrule
Price & 23,768 & 253,646.20 & 170,987.20 & 10,000 & 127,000 & 326,374.8 & 1,000,000 \\ 
Year & 23,768 & 2,004.89 & 6.59 & 1,995 & 1,999 & 2,010 & 2,018 \\ 
ln(volume) & 23,768 & 286.00 & 188.14 & 0.00 & 215.50 & 370.89 & 1,926.45 \\ 
ln(area) & 23,768 & 63.25 & 28.44 & 4.94 & 45.74 & 72.46 & 1,058.70 \\ 
ln(dist. city center) & 23,768 & 2,477.63 & 1,007.55 & 112.78 & 1,652.39 & 3,166.05 & 5,023.62 \\ 
New & 23,768 & 0.04 & 0.20 & 0 & 0 & 0 & 1 \\ 
Type: detached & 23,768 & 0.13 & 0.34 & 0 & 0 & 0 & 1 \\ 
Type: semi-detached & 23,768 & 0.35 & 0.48 & 0 & 0 & 1 & 1 \\ 
Type: terraced & 23,768 & 0.52 & 0.50 & 0 & 0 & 1 & 1 \\ 
Georgian & 23,768 & 0.03 & 0.16 & 0 & 0 & 0 & 1 \\ 
Early Vic. & 23,768 & 0.15 & 0.36 & 0 & 0 & 0 & 1 \\ 
Late Vic./Edw. & 23,768 & 0.21 & 0.40 & 0 & 0 & 0 & 1 \\ 
Interwar & 23,768 & 0.32 & 0.47 & 0 & 0 & 1 & 1 \\ 
Postwar & 23,768 & 0.23 & 0.42 & 0 & 0 & 0 & 1 \\ 
Contemporary & 23,768 & 0.03 & 0.17 & 0 & 0 & 0 & 1 \\ 
Revival & 23,768 & 0.04 & 0.20 & 0 & 0 & 0 & 1 \\ 
Neigh: Georgian & 23,768 & 0.02 & 0.12 & 0 & 0 & 0 & 1 \\ 
Neigh: Early Vic. & 23,768 & 0.14 & 0.35 & 0 & 0 & 0 & 1 \\ 
Neigh: Late V./Edw. & 23,768 & 0.22 & 0.41 & 0 & 0 & 0 & 1 \\ 
Neigh: Interwar & 23,768 & 0.32 & 0.47 & 0 & 0 & 1 & 1 \\ 
Neigh: Postwar & 23,768 & 0.25 & 0.43 & 0 & 0 & 1 & 1 \\ 
Neigh: Contemporary & 23,768 & 0.02 & 0.15 & 0 & 0 & 0 & 1 \\ 
Neigh: Revival & 23,768 & 0.03 & 0.18 & 0 & 0 & 0 & 1 \\ 
\bottomrule
\end{tabular} 
\begin{minipage}{\textwidth}
\vspace{0.25cm}
\singlespacing
\footnotesize \emph{Notes:} Summary statistics for sample of 23,768 residential real estate transactions for the city of Cambridge (UK) between 1995 and 2018 where buildings could be matched with Google Street View images. The buildings floor plate (in $m^2$) is based on OS maps and the buildings' volumes are estimated from digital elevation models, as suggested by @Lindenthal2017a. We control for the location of each building by calculating the distance to the city center proxied by Great St. Mary's Church, and non-parametrically by using 69 location dummies (based on LSOA). Building styles based on machine classifications.
\end{minipage}
\end{table}

@@

Georgian & 0.19 & 0.25 & 0.56** & -- & -- & -- & -- \\ 
   & (0.29) & (0.25) & (0.27) & -- & -- & -- & -- \\ 
Early Vic. & -0.14 & -0.04 & -0.13 & 0.03 & 0.07 & -- & 0.10 \\ 
  & (0.18) & (0.09) & (0.12) & (0.09) & (0.08) & -- & (0.10) \\ 
Late V./Edw. & -0.004 & 0.002 & -0.01 & 0.18* & 0.08 & -- & 0.13 \\ 
   & (0.19) & (0.11) & (0.13) & (0.10) & (0.10) & -- & (0.12) \\ 
Interwar & -0.23 & -0.20** & -0.17 & -0.03 & -0.07 & -- & -0.15 \\ 
  & (0.19) & (0.09) & (0.11) & (0.08) & (0.06) & -- & (0.09) \\ 
Postwar & 0.18 & 0.03 & -0.05 & 0.15** & 0.14*** & -- & 0.01 \\ 
   & (0.28) & (0.11) & (0.12) & (0.07) & (0.05) & -- & (0.09) \\ 
Contemporary & -- & -- & -- & -- & -- & -- & -- \\ 
  & -- & -- & -- & -- & -- & -- & -- \\ 
  Revival & -- & -- & -0.43*** & 0.09 & 0.04 & -- & -0.04 \\ 
   & -- & -- & (0.15) & (0.08) & (0.07) & -- & (0.09) \\ 
\bottomrule
\end{tabular}
\begin{minipage}{\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:} $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01. Standard errors are robust (White's estimator). This table features the coefficients for interaction terms of building style and neighborhood styles only, while Table \ref{tab:hedreg}, Column 8 presents all other coefficients for this model.
\end{minipage}

\end{table}

>>>>

---
title: "Machine Learning, Architectural Styles and Property Values"
author: "Thies Lindenthal* (University of Cambridge) & Erik B. Johnson (University of Alabama)"
date: "`r format(Sys.time(), '%d %B %Y')`"
tags: Externalities of architecture, machine learning
fig_caption: yes
header-includes:
   - \usepackage{soul}
   - \usepackage{color}
   - \sethlcolor{yellow}
   - \usepackage{float}
   - \usepackage{booktabs}
   - \usepackage{graphicx}
   - \usepackage{grffile}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{multirow}
   - \doublespacing
   - \usepackage[backend=bibtex, style=authoryear]{biblatex}

fig_width: 6
fig_height: 4
output: 
  pdf_document:
    keep_tex: true
    number_sections: true
bibliography: references.bib
abstract: "This paper introduces an algorithm that collects pictures of individual buildings from Google Street View. It then trains a deep convolutional neural network (CNN) to classify residential buildings into architectural styles, taking into account spatial dependencies of these styles explicitly. Third, it investigates whether architectural styles influence house prices. For re-sales, the architectural style is a significant determinant of transaction prices while no such effect is found for new buildings. Importantly, this paper explores how to detect and overcome biases in machine learning methods through a large-scale comparison of predictions and expert classifications, partially opening up the `black box' of automatic building classification systems." 
---

\let\svthefootnote\thefootnote
\let\thefootnote\relax\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Peter Brummund, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}
\addtocounter{footnote}{-1}\let\thefootnote\svthefootnote

\newpage

# Introduction

If new developments were only pleasing to the eye, so Britain's *Building Better, Building Beautiful Commission* advising Government, nimbyism would cease and housing supply could finally reach the levels demanded by a growing and more affluent population [@economistlineofbeauty2018]. The inaugural chair of this new commission boldly suggested that if we built *``as our Georgian and Victorian forebears built* [...]. *All objections to new building would slip away in the sheer relief of the public''* [@scruton2018]. Even Prince Charles, in similar spirit, put forward ten principles for urban growth and architecture that emphasize tradition and aesthetics [@princecharles2014]. When governments and princes occupy themselves with beauty, economists surely may take a closer look as well. In this paper, we aim to identify the architectural styles of residential buildings using computer vision techniques and to empirically search for any transaction price premia associated with these styles. If aesthetics were as strong a force in the built environment as claimed, we should indeed find a significant effect on property prices.

@@

Specifically, we first look up the nearest Google Street View panorama point (green dot in Figure \ref{fig:fan}) based on the centroid (red dot) coordinates of a given building obtained from Ordnance Survey maps. We then perform a viewshed analysis and identify which exterior walls are visible from the panorama point, ignoring any wall segments where the direct line of sight from the panorama point is obstructed by other buildings. We can then estimate the camera bearing (green line) and zoom factor, based on the fan of the lines of sight (in blue).\footnote{Code necessary to replicate this analysis is available from the authors' GitHub.com code repositories.}

@@

\newpage

@@

\newpage

\begin{table}[!htbp] \centering 
  \caption{Summary statistics residential property transactions} 
  \label{tab:sumstats} 
\begin{tabular}{@{\extracolsep{5pt}}lccccccc} 
\toprule
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Pctl(25)} & \multicolumn{1}{c}{Pctl(75)} & \multicolumn{1}{c}{Max} \\ 
\midrule
Price & 23,768 & 253,646.20 & 170,987.20 & 10,000 & 127,000 & 326,374.8 & 1,000,000 \\ 
Year & 23,768 & 2,004.89 & 6.59 & 1,995 & 1,999 & 2,010 & 2,018 \\ 
ln(volume) & 23,768 & 286.00 & 188.14 & 0.00 & 215.50 & 370.89 & 1,926.45 \\ 
ln(area) & 23,768 & 63.25 & 28.44 & 4.94 & 45.74 & 72.46 & 1,058.70 \\ 
ln(dist. city center) & 23,768 & 2,477.63 & 1,007.55 & 112.78 & 1,652.39 & 3,166.05 & 5,023.62 \\ 
New & 23,768 & 0.04 & 0.20 & 0 & 0 & 0 & 1 \\ 
Type: detached & 23,768 & 0.13 & 0.34 & 0 & 0 & 0 & 1 \\ 
Type: semi-detached & 23,768 & 0.35 & 0.48 & 0 & 0 & 1 & 1 \\ 
Type: terraced & 23,768 & 0.52 & 0.50 & 0 & 0 & 1 & 1 \\ 
Georgian & 23,768 & 0.03 & 0.16 & 0 & 0 & 0 & 1 \\ 
Early Vic. & 23,768 & 0.15 & 0.36 & 0 & 0 & 0 & 1 \\ 
Late Vic./Edw. & 23,768 & 0.21 & 0.40 & 0 & 0 & 0 & 1 \\ 
Interwar & 23,768 & 0.32 & 0.47 & 0 & 0 & 1 & 1 \\ 
Postwar & 23,768 & 0.23 & 0.42 & 0 & 0 & 0 & 1 \\ 
Contemporary & 23,768 & 0.03 & 0.17 & 0 & 0 & 0 & 1 \\ 
Revival & 23,768 & 0.04 & 0.20 & 0 & 0 & 0 & 1 \\ 
Neigh: Georgian & 23,768 & 0.02 & 0.12 & 0 & 0 & 0 & 1 \\ 
Neigh: Early Vic. & 23,768 & 0.14 & 0.35 & 0 & 0 & 0 & 1 \\ 
Neigh: Late V./Edw. & 23,768 & 0.22 & 0.41 & 0 & 0 & 0 & 1 \\ 
Neigh: Interwar & 23,768 & 0.32 & 0.47 & 0 & 0 & 1 & 1 \\ 
Neigh: Postwar & 23,768 & 0.25 & 0.43 & 0 & 0 & 1 & 1 \\ 
Neigh: Contemporary & 23,768 & 0.02 & 0.15 & 0 & 0 & 0 & 1 \\ 
Neigh: Revival & 23,768 & 0.03 & 0.18 & 0 & 0 & 0 & 1 \\ 
\bottomrule
\end{tabular} 
\begin{minipage}{\textwidth}
\vspace{0.25cm}
\singlespacing
\footnotesize \emph{Notes:} Summary statistics for sample of 23,768 residential real estate transactions for the city of Cambridge (UK) between 1995 and 2018 where buildings could be matched with Google Street View images. The buildings floor plate (in $m^2$) is based on OS maps and the buildings' volumes are estimated from digital elevation models (Lindenthal, 2017b). We control for the location of each building by calculating the distance to the city center proxied by Great St. Mary's Church, and non-parametrically by using 69 location dummies (based on LSOA). Building styles based on machine classifications.
\end{minipage}
\end{table}

\newpage

@@

Georgian & 0.19 & 0.25 & 0.56** & -- & -- & -- & -- \\ 
   & (0.29) & (0.25) & (0.27) & -- & -- & -- & -- \\ 
Early Vic. & -0.14 & -0.04 & -0.13 & 0.03 & 0.07 & -- & 0.10 \\ 
  & (0.18) & (0.09) & (0.12) & (0.09) & (0.08) & -- & (0.10) \\ 
Late V./Edw. & -0.004 & 0.002 & -0.01 & 0.18* & 0.08 & -- & 0.13 \\ 
   & (0.19) & (0.11) & (0.13) & (0.10) & (0.10) & -- & (0.12) \\ 
Interwar & -0.23 & -0.20** & -0.17 & -0.03 & -0.07 & -- & -0.15 \\ 
  & (0.19) & (0.09) & (0.11) & (0.08) & (0.06) & -- & (0.09) \\ 
Postwar & 0.18 & 0.03 & -0.05 & 0.15** & 0.14*** & -- & 0.01 \\ 
   & (0.28) & (0.11) & (0.12) & (0.07) & (0.05) & -- & (0.09) \\ 
Contemporary & -- & -- & -- & -- & -- & -- & -- \\ 
  & -- & -- & -- & -- & -- & -- & -- \\ 
  Revival & -- & -- & -0.43*** & 0.09 & 0.04 & -- & -0.04 \\ 
   & -- & -- & (0.15) & (0.08) & (0.07) & -- & (0.09) \\ 
\bottomrule
\end{tabular}
\begin{minipage}{\textwidth}
\vspace{0.25cm}
\footnotesize \emph{Notes:} $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01. Standard errors are robust (White's estimator). This table features the coefficients for interaction terms of building style and neighborhood styles only, while Table \ref{tab:hedreg}, Column 8 presents all other coefficients for this model.
\end{minipage}
\end{table}

\newpage

### 2020-04-13 text/manuscript_assa.tex
Tables and Figures on separate pages
Are we back in the game?

  \title{Machine Learning, Architectural Styles and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{14 September 2019}

@@

\begin{document}
\maketitle
\begin{abstract}
This paper first introduces an algorithm that collects pictures of
individual buildings from Google Street View. Second, it trains a deep
convolutional neural network (CNN) to classify residential buildings
into architectural styles, taking into account spatial dependencies of
these styles. Third, it investigates whether architectural styles
influence house prices. For re-sales, the architectural style is a
significant determinant of transaction prices while no such effect is
found for new buildings. Additionally, we are able to provide guidance
on how to detect and overcome some of the limitations of machine
learning methods through a large-scale comparison of predictions and
expert classifications.
\end{abstract}

@@

The aesthetics of different architectural styles make for an easy
cocktail party conversation topic -- and they may also influence market
prices. Using architectural assessments by human experts, Buitelaar and
Schilder (2017) find a sizable premium of 5\% for new buildings in the
Netherlands that refer to traditional styles and a staggering 15\%
premium for new buildings that closely follow traditional shapes, facade
composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location, or year of construction.
These controls are crucial, as earlier work has established that age and
style variables tend to be highly correlated. Coulson and McMillen
(2008), for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary styles (vis-à-vis more historic styles). Francke and Minne
(2017) investigate the depreciation of residential real estate in the
Netherlands and decompose land versus structure values singling out the
effect of ``physical deterioration, functional obsolescence and and
vintage effects''. They find that buildings from the 1930s carry a
strong price premium.

@@

Unfortunately, the traditional approaches chosen by e.g. Buitelaar and
Schilder (2017) or Ahlfeldt and Mastro (2012) do not scale well. Each
observation has to be classified into architectural styles by a human
expert, which is time-consuming and costly, imposing an upper limit on
the number of observations and the level of detail captured for each
observation in any given sample. Google Street View offers a solution to
this nexus as it captures the images of almost all buildings in many
cities around the world at high level of accuracy and detail. The
challenge we address in this paper is to extract and utilize building
level information from this ubiquitous sensor, using freely available
deep learning techniques.

@@

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} style while starter homes are disproportionally
more likely to be classified as e.g. \emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for machine's classifications
versus the estimates for architects' classifications in the subsequent
hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

@@

The large ground truth data set allows for a comparison of human expert
versus machine classifications. Poor image quality, for instance
obstructed views or the lack of informative features such as windows, is
correlated with misclassifications. Importantly, for cases in which the
ML classifiers are relatively indecisive between classes, visual cues
that are correlated with the exterior's style but not part of a
building's architecture (for instance trees or cars) bias the estimates.
Fore these less certain cases, more valuable homes are systematically
more likely to be classified as e.g. \emph{Late Victorian/Edw.} and less
likely to be considered \emph{Contemporary}. A follow up study could
investigate which elements in the building pictures lead to the
misclassifications (Ribeiro, Singh, and Guestrin 2016). Given the danger
of systematic biases, one should remain wary of ML estimates derived
from data sets sizable enough to train models -- but too small to
investigate any biases.

@@

\leavevmode\hypertarget{ref-Helbich2013}{}%
Helbich, Marco, Andreas Jochem, Werner Mücke, and Bernhard Höfle. 2013.
``Boosting the predictive accuracy of urban hedonic house price models
through airborne laser scanning.'' \emph{Computers, Environment and
Urban Systems} 39. Elsevier Ltd: 81--92.
\url{https://doi.org/10.1016/j.compenvurbsys.2013.01.001}.

@@

\leavevmode\hypertarget{ref-Liu2017}{}%
Liu, Lun, Elisabete A. Silva, Chunyang Wu, and Hui Wang. 2017. ``A
machine learning-based method for the large-scale evaluation of the
qualities of the urban environment.'' \emph{Computers, Environment and
Urban Systems} 65. The Authors: 113--25.
\url{https://doi.org/10.1016/j.compenvurbsys.2017.06.003}.

@@

\leavevmode\hypertarget{ref-economistlineofbeauty2018}{}%
The Economist. 2018. ``The line of beauty.'' \emph{The Economist},
November. London.
\url{https://www.economist.com/britain/2018/11/17/how-to-defeat-nimbyism-build-more-beautiful-houses}.

>>>>

  \title{Machine Learning, Architectural Styles and Property Values}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{13 April 2020}

@@

\begin{document}
\maketitle
\begin{abstract}
This paper introduces an algorithm that collects pictures of individual
buildings from Google Street View. It then trains a deep convolutional
neural network (CNN) to classify residential buildings into
architectural styles, taking into account spatial dependencies of these
styles explicitly. Third, it investigates whether architectural styles
influence house prices. For re-sales, the architectural style is a
significant determinant of transaction prices while no such effect is
found for new buildings. Importantly, this paper explores how to detect
and overcome biases in machine learning methods through a large-scale
comparison of predictions and expert classifications, partially opening
up the `black box' of automatic building classification systems.
\end{abstract}

@@

The aesthetics of different architectural styles make for an easy
cocktail party conversation topic -- and they may also influence market
prices. Using architectural assessments by human experts, Buitelaar and
Schilder (2017) find a sizable premium of 5\% for new buildings in the
Netherlands that refer to traditional styles and a staggering 15\%
premium for new buildings that closely follow traditional shapes, facade
composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location, or year of construction.
These controls are crucial, as earlier work has established that age and
style variables tend to be highly correlated. Coulson and McMillen
(2008), for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary styles (vis-à-vis more historic styles). Francke and van de
Minne (2017) investigate the depreciation of residential real estate in
the Netherlands and decompose land versus structure values singling out
the effect of ``physical deterioration, functional obsolescence and and
vintage effects''. They find that buildings from the 1930s carry a
strong price premium.

@@

Unfortunately, the traditional approaches chosen by e.g.~Buitelaar and
Schilder (2017) or Ahlfeldt and Mastro (2012) do not scale well. Each
observation has to be classified into architectural styles by a human
expert, which is time-consuming and costly, imposing an upper limit on
the number of observations and the level of detail captured for each
observation in any given sample. Google Street View offers a solution to
this nexus as it captures the images of almost all buildings in many
cities around the world at high level of accuracy and detail. The
challenge we address in this paper is to extract and utilize building
level information from this ubiquitous sensor, using freely available
deep learning techniques.

@@

Will the resulting classifiers be biased with respect to property
values, picking up correlated but irrelevant cues such as greenery,
upkeep, or cars brands instead of building exteriors? Are, for instance,
more valuable houses more likely to be classified as the popular
\emph{Late Victorian} style while starter homes are disproportionally
more likely to be classified as e.g.~\emph{Early Victorian}? We expect
that in cases in which the ensemble provided a high confidence score,
building features are clearly detectable on the pictures. For close
calls, however, where models within the ensemble cannot agree, other
factors might tilt the scale towards an incorrect class. A sizable
difference in estimated marginal prices for machine's classifications
versus the estimates for architects' classifications in the subsequent
hedonic regression (Eq. \ref{eq:hed}, below) would indicate such a bias.

@@

The large ground truth data set allows for a comparison of human expert
versus machine classifications. Poor image quality, for instance
obstructed views or the lack of informative features such as windows, is
correlated with misclassifications. Importantly, for cases in which the
ML classifiers are relatively indecisive between classes, visual cues
that are correlated with the exterior's style but not part of a
building's architecture (for instance trees or cars) bias the estimates.
Fore these less certain cases, more valuable homes are systematically
more likely to be classified as e.g.~\emph{Late Victorian/Edw.} and less
likely to be considered \emph{Contemporary}. A follow up study could
investigate which elements in the building pictures lead to the
misclassifications (Ribeiro, Singh, and Guestrin 2016). Given the danger
of systematic biases, one should remain wary of ML estimates derived
from data sets sizable enough to train models -- but too small to
investigate any biases.

@@

\leavevmode\hypertarget{ref-Helbich2013}{}%
Helbich, Marco, Andreas Jochem, Werner Mücke, and Bernhard Höfle. 2013.
``Boosting the predictive accuracy of urban hedonic house price models
through airborne laser scanning.'' \emph{Computers, Environment and
Urban Systems} 39: 81--92.
\url{https://doi.org/10.1016/j.compenvurbsys.2013.01.001}.

@@

\leavevmode\hypertarget{ref-Liu2017}{}%
Liu, Lun, Elisabete A. Silva, Chunyang Wu, and Hui Wang. 2017. ``A
machine learning-based method for the large-scale evaluation of the
qualities of the urban environment.'' \emph{Computers, Environment and
Urban Systems} 65: 113--25.
\url{https://doi.org/10.1016/j.compenvurbsys.2017.06.003}.

@@

\leavevmode\hypertarget{ref-economistlineofbeauty2018}{}%
The Economist. 2018. ``The line of beauty.'' \emph{The Economist},
November.
\url{https://www.economist.com/britain/2018/11/17/how-to-defeat-nimbyism-build-more-beautiful-houses}.
